---
title: "Graphs"
author: "Claire Herdeman"
date: "8/15/2018"
output: 
  html_document:
    code_folding: hide
    toc: true
    toc_depth: 2
    toc_float:
      collapsed: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r dependencies, include = F}
library(knitr)
library(meta)
library(dplyr)
library(ggplot2)
```

# Define Functions {.tabset}

## Barplots 

```{r}
# Note: this is a simplified version of an internal AI graphing function
"This function wraps ggplot2's geom_bar plot to more easily create replicable barplots with a similar style."
barplot_simple <- function(df, x, y,
                     fillVar = NULL, fillColor = NULL,
                     barwidth = 0.9,
                     upper.err = 'upper.err', lower.err = 'lower.err',
                     hasErrBar = TRUE,
                     errorbarwidth = 0.15, errorbarcolor = '#575757',
                     xlabel = NULL,
                     yLabelPlace = NULL,
                     yLabelName = 'ylabel',
                     xtitle = '', ytitle = '',
                     hasLegend = FALSE,
                     lab_size = 4.75,
                     ncol = NULL,
                     nrow = NULL){
  require(ggplot2)

  # Make Plots -----
  dodge <- position_dodge(width = barwidth)

  # Base Plot
  pp <- ggplot(df, aes_string(x = x, y = y, fill = fillVar)) +
    geom_bar(stat = "identity", width = barwidth, position = dodge)

  # Add Error Bars
  if(hasErrBar){
    pp <- pp + geom_errorbar(aes_string(ymax = upper.err, ymin = lower.err),
                  position = dodge, width = errorbarwidth,
                  colour = errorbarcolor)
  }
  
  # Add Fill Color
  if(!is.null(fillColor)){
    pp <- pp + scale_fill_manual(values = as.character(df[[fillColor]]))
  }
  
  # Add Axes
  pp <- pp + labs(x = xtitle, y = ytitle)

  # Remove Legend
  if(hasLegend == F){
    pp <- pp + guides(fill=FALSE)
  }

  return(pp)
}
```

## Percentage Point Transformation

```{r}
# This is a version of AI's transform function
"This function converts standard model outputs from decimals to percentage points."
transform <- function(df, var_estimate, var_se,
                         var_ci_lower = NULL, var_ci_upper = NULL){

  df[[var_estimate]] <- df[[var_estimate]] * 100
  df[[var_se]] <- df[[var_se]] * 100

  if(!is.null(var_ci_lower))
    df[[var_ci_lower]] <- df[[var_ci_lower]] * 100
  if(!is.null(var_ci_upper))
    df[[var_ci_upper]] <- df[[var_ci_upper]] * 100

  df
}
```

# Load Data

```{r}
# Working Directory
wd <- "~/projects/ai_code_sample/Meta-Analysis/data/"

# Read in files
original <- read.csv(paste0(wd, "old_data.csv"))
update <- read.csv(paste0(wd, "new_data.csv"))

# Merge 
df <- rbind(original, update)
```

# Validate Prior Results {.tabset}

The purpose of this section is to recreate the graphs from the original analysis to validate that the updated methodology (in R) creates the same output as the previous analysis (in Stata).

## Gender

```{r}
g1 <- metagen(TE=treatment_effect, seTE=standard_error, n.c=control_n, n.e = treat_n, data=original, subset=group == "Male")
g2 <- metagen(TE=treatment_effect, seTE=standard_error, n.c=control_n, n.e = treat_n, data=original, subset=group == "Female")

gender_old <- data.frame('label'=c('Male', 'Female'), 'te'=c(g1$TE.random, g2$TE.random),
                         'se' = c(g1$seTE.random, g2$seTE.random),
                         'lower.err'=c(g1$lower.random, g2$lower.random), 
                         'upper.err'=c(g1$upper.random, g2$upper.random))

gender_old$color <- c("light blue", "gray")

gender_old <- transform(gender_old, 'te', 'se', 'lower.err', 'upper.err')

p1 <- barplot_simple(gender_old, 'label', 'te', fillVar='label', fillColor = 'color',
                         yLabelPlace = 'offset',
                         yLabelName = round(gender_old$te, digits = 1),
                         ytitle = 'Treatment Effect (percentage points)',
                         xtitle = 'Gender')

p1
```

## Age

```{r}
# Age
a1 <- metagen(TE=treatment_effect, seTE=standard_error, n.c=control_n, n.e = treat_n, data=original, subset=group == "18-24")
a2 <- metagen(TE=treatment_effect, seTE=standard_error, n.c=control_n, n.e = treat_n, data=original, subset=group == "25-30")
a3 <- metagen(TE=treatment_effect, seTE=standard_error, n.c=control_n, n.e = treat_n, data=original, subset=group == "31-45")
a4 <- metagen(TE=treatment_effect, seTE=standard_error, n.c=control_n, n.e = treat_n, data=original, subset=group == "46-60")
a5 <- metagen(TE=treatment_effect, seTE=standard_error, n.c=control_n, n.e = treat_n, data=original, subset=group == "61+")

age_old <- data.frame('label'=c('18-24', '25-30', '31-45', '46-60', '61+'), 
                      'te'=c(a1$TE.random, a2$TE.random, a3$TE.random, a4$TE.random, a5$TE.random),
                       'se' = c(a1$seTE.random, a2$seTE.random, a3$seTE.random, a4$seTE.random, a5$seTE.random),
                       'lower.err'=c(a1$lower.random, a2$lower.random, a3$lower.random, a4$lower.random, a5$lower.random), 
                       'upper.err'=c(a1$upper.random, a2$upper.random, a3$upper.random, a4$upper.random, a5$upper.random))

age_old$colors <- c("gray", "light blue", "light green", "red", "orange")

age_old <- transform(age_old, 'te', 'se', 'lower.err', 'upper.err')

p2 <- barplot_simple(age_old, 'label', 'te', fillVar='label', fillColor = 'colors', 
                         yLabelPlace = 'offset',
                         yLabelName = round(age_old$te, digits = 1),
                         ytitle = 'Treatment Effect (percentage points)',
                         xtitle = 'Age')

p2
```

## Race

```{r}
# Race
r1 <- metagen(TE=treatment_effect, seTE=standard_error, n.c=control_n, n.e = treat_n, data=original, subset=group == "White")
r2 <- metagen(TE=treatment_effect, seTE=standard_error, n.c=control_n, n.e = treat_n, data=original, subset=group == "Black")
r3 <- metagen(TE=treatment_effect, seTE=standard_error, n.c=control_n, n.e = treat_n, data=original, subset=group == "Latinx")

race_old <- data.frame('label'=c('White', 'African American', 'Latinx'), 
                         'te'=c(r1$TE.random, r2$TE.random, r3$TE.random),
                         'se' = c(r1$seTE.random, r2$seTE.random, r3$seTE.random),
                         'lower.err'=c(r1$lower.random, r2$lower.random, r3$lower.random), 
                         'upper.err'=c(r1$upper.random, r2$upper.random, r3$upper.random))

race_old$colors <- c("gray", "light blue", "light green")

race_old <- transform(race_old, 'te', 'se', 'lower.err', 'upper.err')

p3 <- barplot_simple(race_old, 'label', 'te', fillVar='label', fillColor = 'colors', 
                         yLabelPlace = 'offset',
                         yLabelName = round(race_old$te, digits = 1),
                         ytitle = 'Treatment Effect (percentage points)',
                         xtitle = 'Race')

p3
```

# Updated Results {.tabset}

The updated results, including additional experiments, can be compared side by side with the results of the prior analysis. As the data is randomly generated, no meaningful conclusions can be drawn from this comparison.

## Gender

```{r}
tg1 <- metagen(TE=treatment_effect, seTE=standard_error, n.c=control_n, n.e = treat_n, data=df, subset=group == "Male")
tg2 <- metagen(TE=treatment_effect, seTE=standard_error, n.c=control_n, n.e = treat_n, data=df, subset=group == "Female")

gender_new <- data.frame('label'=c('Male', 'Female'), 'te'=c(tg1$TE.random, tg2$TE.random),
                         'se' = c(tg1$seTE.random, tg2$seTE.random),
                         'lower.err'=c(tg1$lower.random, tg2$lower.random), 
                         'upper.err'=c(tg1$upper.random, tg2$upper.random))

gender_new$color <- c("light blue", "gray")

gender_new <- transform(gender_new, 'te', 'se', 'lower.err', 'upper.err')

tp1 <- barplot_simple(gender_new, 'label', 'te', fillVar='label', fillColor = 'color',
                         yLabelPlace = 'offset',
                         yLabelName = round(gender_old$te, digits = 1),
                         ytitle = 'Treatment Effect (percentage points)',
                         xtitle = 'Gender')

tp1
```

## Age

```{r}
# Age
ta1 <- metagen(TE=treatment_effect, seTE=standard_error, n.c=control_n, n.e = treat_n, data=df, subset=group == "18-24")
ta2 <- metagen(TE=treatment_effect, seTE=standard_error, n.c=control_n, n.e = treat_n, data=df, subset=group == "25-30")
ta3 <- metagen(TE=treatment_effect, seTE=standard_error, n.c=control_n, n.e = treat_n, data=df, subset=group == "31-45")
ta4 <- metagen(TE=treatment_effect, seTE=standard_error, n.c=control_n, n.e = treat_n, data=df, subset=group == "46-60")
ta5 <- metagen(TE=treatment_effect, seTE=standard_error, n.c=control_n, n.e = treat_n, data=df, subset=group == "61+")

age_new <- data.frame('label'=c('18-24', '25-30', '31-45', '46-60', '61+'), 
                      'te'=c(ta1$TE.random, ta2$TE.random, ta3$TE.random, ta4$TE.random, ta5$TE.random),
                       'se' = c(ta1$seTE.random, ta2$seTE.random, ta3$seTE.random, ta4$seTE.random, ta5$seTE.random),
                       'lower.err'=c(ta1$lower.random, ta2$lower.random, ta3$lower.random, ta4$lower.random, ta5$lower.random), 
                       'upper.err'=c(ta1$upper.random, ta2$upper.random, ta3$upper.random, ta4$upper.random, ta5$upper.random))

age_new$colors <- c("gray", "light blue", "light green", "red", "orange")

age_new <- transform(age_new, 'te', 'se', 'lower.err', 'upper.err')

t2 <- barplot_simple(age_new, 'label', 'te', fillVar='label', fillColor = 'colors', 
                         yLabelPlace = 'offset',
                         yLabelName = round(age_old$te, digits = 1),
                         ytitle = 'Treatment Effect (percentage points)',
                         xtitle = 'Age')

t2
```

## Race

```{r}
# Race
tr1 <- metagen(TE=treatment_effect, seTE=standard_error, n.c=control_n, n.e = treat_n, data=df, subset=group == "White")
tr2 <- metagen(TE=treatment_effect, seTE=standard_error, n.c=control_n, n.e = treat_n, data=df, subset=group == "Black")
tr3 <- metagen(TE=treatment_effect, seTE=standard_error, n.c=control_n, n.e = treat_n, data=df, subset=group == "Latinx")

race_new <- data.frame('label'=c('White', 'African American', 'Latinx'), 
                         'te'=c(tr1$TE.random, tr2$TE.random, tr3$TE.random),
                         'se' = c(tr1$seTE.random, tr2$seTE.random, tr3$seTE.random),
                         'lower.err'=c(tr1$lower.random, tr2$lower.random, tr3$lower.random), 
                         'upper.err'=c(tr1$upper.random, tr2$upper.random, tr3$upper.random))

race_new$colors <- c("gray", "light blue", "light green")

race_new <- transform(race_new, 'te', 'se', 'lower.err', 'upper.err')

tp3 <- barplot_simple(race_new, 'label', 'te', fillVar='label', fillColor = 'colors', 
                         yLabelPlace = 'offset',
                         yLabelName = round(race_old$te, digits = 1),
                         ytitle = 'Treatment Effect (percentage points)',
                         xtitle = 'Race')

tp3
```